---
title: "JSOES Bongo Data Exploratory Analysis"
author: "Markus Min"
date: "`r Sys.Date()`"
output:
  html_document
---
```{r, echo = FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

This page contains exploratory analysis for the Bongo net data from [JSOES](https://markusmin.github.io/JSOES-SDM/jsoes_survey_description.html). Please note that because the Bongo net tows occur during daylight hours, the tow composition is biased towards taxa/life stages that do not vertically migrate. While different life stages of some species are counted separately for this data, for this exploratory analysis we summed abundances across life stages. 

## Data Overview

We will first explore the most common and most abundant taxa in this dataset.


```{r load_libraries_data}
library(readxl)
library(here)
library(janitor)
library(tidyverse)
library(kableExtra)
library(sf)
library(gstat)
library(sp)
library(ggthemes)
library(ape)


bongo <- clean_names(read_excel(here::here("Data", "Markus_Min_Plankton Density for Trophic Summary Query_10.11.24.xlsx")))
# Drop any samples that don't have spatial data
filter(bongo, !(is.na(dec_long))) -> bongo
# this drops one sample: 062904NH30 (NH30 in June 2004)

# turn this into an sf object
st_as_sf(bongo, coords = c("dec_long", "dec_lat"), crs = 4326) -> bongo_sf

# change CRS to UTM zone 10 (to work in meters)
UTM_zone_10_crs <- 32610
bongo_sf_proj <- sf::st_transform(bongo_sf, crs = UTM_zone_10_crs)

# make this projection into kilometers to help with interpretability
bongo_sf_proj_km <- st_as_sf(bongo_sf_proj$geometry/1000, crs = UTM_zone_10_crs)

# extract geometry
as.data.frame(st_coordinates(bongo_sf_proj_km)) -> bongo_km
# add this back to bongo (X and Y now represent eastings and northings in km)
bind_cols(bongo, bongo_km) -> bongo

trophic_groupings <- clean_names(read_excel(here::here("Data", "Markus_Min_Trophic Groupings_10.2.24.xlsx")))

# change genus species to be consistent across these two files, and usual 
# latin name and common name capitalization
bongo$genus_species <- str_to_sentence(bongo$genus_species)
trophic_groupings$genus_species <- str_to_sentence(trophic_groupings$genus_species)
trophic_groupings$common_name <- str_to_title(trophic_groupings$common_name)

trophic_groupings %>% 
  dplyr::select(genus_species, common_name) %>% 
  filter(!(duplicated(genus_species))) -> common_name_df

# get all unique tows
bongo %>% 
  mutate(sampleID = paste0(station, "_", sample_date)) %>% 
  relocate(sampleID) %>% 
  dplyr::select(-c(genus_species, life_history_stage, id_code, count, sum_of_density_number_m3)) %>% 
  filter(!(duplicated(sampleID))) -> bongo_tows

# complete bongo data, so that we have zeros for each taxon in tows where they weren't caught
bongo %>% 
  mutate(sampleID = paste0(station, "_", sample_date)) %>% 
  relocate(sampleID) %>% 
  # drop tow information; we will add this back later
  dplyr::select(-c(id_code, cruise_number, month, year, station_code, net_type, sample_date,
           sample_time, dec_lat, dec_long, station_depth_m, transect_name, station_number, 
           station, n_cr_s, day_night, volume_filtered_m3)) %>% 
  complete(sampleID, 
           nesting(genus_species, life_history_stage), 
           fill = list(count = 0, sum_of_density_number_m3 = 0)) -> bongo_catch

bongo_catch %>% 
  left_join(bongo_tows, by = "sampleID") -> bongo_catch





# add common name
bongo_catch %>% left_join(common_name_df, by = "genus_species") -> bongo_catch

# create a version of this where we collapse all life history stages together
bongo_catch %>% 
  dplyr::select(sampleID, genus_species, count, sum_of_density_number_m3) %>% 
  group_by(sampleID, genus_species) %>% 
  summarise_if(is.numeric, sum) -> bongo_catch_LH_collapse

bongo_catch_LH_collapse %>% 
  left_join(bongo_tows, by = "sampleID") %>% 
  ungroup() -> bongo_catch_LH_collapse
```

```{r load_spatial_data, cache = TRUE}
#### Spatial data ####
usa_spdf <- st_read("/Users/markusmin/Documents/ESA_RF_2021/map_files/USA_adm0.shp", quiet = TRUE)
# load BC
CAN_spdf <- st_read("/Users/markusmin/Documents/ESA_RF_2021/map_files/canada/lpr_000b16a_e.shp", quiet = TRUE)
BC_spdf <- filter(CAN_spdf, PRENAME == "British Columbia")
BC_proj <- st_transform(BC_spdf, crs = 4326)


# crop them to our desired area
US_west_coast <- sf::st_crop(usa_spdf,
    c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))

BC_coast <- sf::st_crop(BC_proj,
    c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))



# convert both shapefiles to a different projection (UTM zone 10) so that they can be plotted with the sdmTMB output
UTM_zone_10_crs <- 32610

US_west_coast_proj <- sf::st_transform(US_west_coast, crs = UTM_zone_10_crs)
BC_coast_proj <- sf::st_transform(BC_coast, crs = UTM_zone_10_crs)

# make this projection into kilometers
US_west_coast_proj_km <- st_as_sf(US_west_coast_proj$geometry/1000, crs = UTM_zone_10_crs)
BC_coast_proj_km <- st_as_sf(BC_coast_proj$geometry/1000, crs = UTM_zone_10_crs)
```



```{r common_taxa}
## total density
bongo_catch %>% 
  group_by(genus_species) %>% 
  summarise(mean_density_per_m3 = mean(sum_of_density_number_m3),
            sd_density_per_m3 = sd(sum_of_density_number_m3)) %>% 
  arrange(desc(mean_density_per_m3)) %>% 
  left_join(common_name_df, by = "genus_species") %>% 
  relocate(common_name, .after = "genus_species") -> bongo_catch_density_sums

bongo_catch_density_sums[1:10,] %>% 
  mutate(mean_density_per_m3 = round(mean_density_per_m3, 1),
         sd_density_per_m3 = round(sd_density_per_m3, 1)) %>% 
  kbl(caption = "Top ten taxa by mean density.") %>% 
  kable_styling(latex_options = "HOLD_position")

## frequency of occurrence

# get count of each genus_species
bongo_catch_LH_collapse %>% 
  filter(sum_of_density_number_m3 > 0) %>% 
  count(genus_species) %>% 
  arrange(desc(n)) %>% 
  mutate(prop_samples = n/nrow(bongo_tows)) %>% 
  left_join(common_name_df, by = "genus_species") %>% 
  relocate(common_name, .after = "genus_species") -> bongo_freq_occ

bongo_freq_occ[1:10, c("genus_species", "common_name", "prop_samples")] %>% 
  mutate(prop_samples = round(prop_samples, 2)) %>% 
  kbl(caption = "Top ten taxa by frequency of occurrence.") %>% 
  kable_styling(latex_options = "HOLD_position")

```


```{r create_basemap, cache = TRUE}
# create base map
survey_area_basemap_km <- ggplot(US_west_coast_proj_km) +
  geom_sf() +
  geom_sf(data = BC_coast_proj_km) +
  # coord_sf(ylim = c(4922.052, 5342.052), xlim = c(334.8638, 404.8638)) +
  # coord_sf(ylim = c(44,48.5),  xlim = c(-126, -123)) +
  ylab("Latitude")+
  xlab("Longitude")+
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white", color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.14, 0.2),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        axis.ticks = element_blank(),
        axis.text = element_blank())
  # temporary fix for lat/long on axes to just get rid of them - start here to actually fix: https://forum.posit.co/t/converting-axes-to-lat-lon/27181

survey_area_basemap <- ggplot(US_west_coast) +
  geom_sf() +
  geom_sf(data = BC_coast) +
  coord_sf(ylim = c(44,48.5),  xlim = c(-126, -123)) +
  scale_x_continuous(breaks = c(124,125,126)) +
  ylab("Latitude")+
  xlab("Longitude")+
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white", color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.14, 0.2),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))
```


```{r plot-taxon-distribution}
plot_distribution <- function(data, taxon_name){
  # create facet_wrap plot for distribution across all years
  data %>% 
    mutate(encounter = ifelse(sum_of_density_number_m3 == 0, "zero", "non-zero")) %>% 
    dplyr::rename("density (N/m3)" = sum_of_density_number_m3) -> data
  
  survey_area_basemap +
      geom_point(data = data, aes(x = dec_long, y = dec_lat, size = `density (N/m3)`, color = encounter),
                 alpha = 0.5) +
    scale_color_manual(values = c("zero" = "#fc9272", "non-zero" = "#2ca25f")) +
    facet_wrap(~year, nrow = 3) +
    ggtitle(taxon_name) +
    theme(legend.position = c(0.9, 0.15),
        legend.key.height = unit(0.35, "cm"),
        legend.key.width = unit(0.25, "cm"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.spacing.y = unit(0.01, 'cm')) -> species_distribution_plot
  
  return(species_distribution_plot)
}
```
## Annual time series

This Shiny app can be used to explore the abundances of different taxa across the full length of the time series. In this Shiny app, I take the mean log density across the survey region to create a simple index of abundance.

```{r}
knitr::include_app(url = "https://markusmin.shinyapps.io/jsoes_bongo_annual_time_series/")
# <iframe src="https://markusmin.shinyapps.io/jsoes_bongo_annual_time_series/" style="border:none;width:1000px;height:500px;"></iframe>
```


## Plotting distributions of some common taxa

I am currently developing a Shiny app that will allow you to visualize the distributions of different taxa. For now, I provide static maps of two focal taxa for this survey:  *Calanus marshallae* (an abundant cold-water copepod) and *Calanus pacificus* (an abundant warm-water coepod). To demonstrate how we can visualize the abundance of these two key copepod species in space and time, we plot their density in survey catch below.

```{r map_calanus, cache = TRUE}
cama_dist_plot <- plot_distribution(data = subset(bongo_catch_LH_collapse, genus_species == "Calanus marshallae"),
                                    taxon_name = "Calanus marshallae")


ggsave(here::here("docs", "figures", "jsoes_bongo_exploratory", "cama_dist_plot.png"), plot = cama_dist_plot, height = 12, width = 17)


capa_dist_plot <- plot_distribution(data = subset(bongo_catch_LH_collapse, genus_species == "Calanus pacificus"),
                                    taxon_name = "Calanus pacificus")


ggsave(here::here("docs", "figures", "jsoes_bongo_exploratory", "capa_dist_plot.png"), plot = capa_dist_plot, height = 12, width = 17)

# print maps for website
cama_dist_plot
capa_dist_plot

```

```{r calanus_time_series}
create_time_series <- function(data, taxon){
  data %>% 
    mutate(log_density = log(sum_of_density_number_m3 + 1)) %>% 
    group_by(year) %>% 
    summarise(mean_density = mean(log_density),
              sd_density = sd(log_density)) %>% 
    # mutate(CI_95_upper = mean_density + 1.96*sd_density) %>% 
    # mutate(CI_95_lower = mean_density - 1.96*sd_density) %>% 
    mutate(taxon = taxon) -> ts
  
  return(ts)
}

create_time_series(data = subset(bongo_catch_LH_collapse, genus_species == "Calanus marshallae"),
                                 taxon = "Calanus marshallae") -> cama_ts
create_time_series(data = subset(bongo_catch_LH_collapse, genus_species == "Calanus pacificus"),
                   taxon = "Calanus pacificus") -> capa_ts

cama_ts %>% 
  bind_rows(., capa_ts) -> calanus_ts
```

<br>


## Temporal and Spatial Autocorrelation

Before fitting any spatiotemporal models, we must explore the spatial and temporal autocorrelation in the data.

```{r create_station_grid}
# Determine the distance between different stations
# NH30 was apparently only sampled on 2004-06-29; so ignore that one. It's also
# the only station with missing lat/lon info
bongo_tows %>% 
  filter(!(is.na(dec_lat))) %>% 
  filter(!(duplicated(station))) %>% 
  dplyr::select(dec_lat, dec_long, station_depth_m, transect_name, transect_name,
                station_number, station) %>% 
  relocate(station, transect_name, station_number) -> bongo_stations

# convert to sf object
st_as_sf(bongo_stations, coords = c("dec_long", "dec_lat"), crs = 4326) -> bongo_stations_sf

as.data.frame(st_distance(bongo_stations_sf)) -> bongo_stations_distance_matrix
colnames(bongo_stations_distance_matrix) <- bongo_stations$station
rownames(bongo_stations_distance_matrix) <- bongo_stations$station
```

<br>

### Temporal structure

We can first inspect the autocorrelation in our *Calanus marshallae* and *Calanus pacificus* mean annual time series.

```{r coastwide_temporal_autocorrelation}
# It would make sense to try different spatial structures and then test autocorrelation.
# 1) Coastwide; 2) By transect; 3) By station

acf(cama_ts$mean_density, main = "Calanus marshallae, mean log density")
acf(capa_ts$mean_density, main = "Calanus pacificus, mean log density")
```

At the coastwide scale the *Calanus marshallae* time series does not show temporal autocorrelation, but the *Calanus pacificus* time series does at a lag of one year.

I also investigated temporal autocorrelation at the scale of individual stations, and found little evidence for temporal autocorrelation at this scale for these two taxa.

<br>

### Spatial structure

To investigate spatial autocorrelation, I calculated a metric known as Moran's I. Moran's I is a measure of the overall clustering of the spatial data and tests if there is support to reject the null hypothesis of no spatial structure. Given that we did not see much evidence for temporal autocorrelation, I calculated Moran's I separately for each year. To summarize the Moran's I results, I show the p-value for Moran's I for each year, with the blue dashed line showing a p-value of 0.05.

```{r spatial_autocorrelation_functions}
# If you wanted to add a best fit line:
# A best fit line is then plotted through the points.
# For each year, we will plot the best fit semivariogram model, which we selected from a set of five candidate functional response shapes: 1) Exponential, 2) Spherical, 3) Gaussian, 4) Linear, and 5) Nugget only (y-intercept only).



# Set the models to test
# To see the shape of these models, run `show.vgms()`
model.choices <- vgm(model = c("Nug", "Exp", "Sph", "Gau", "Lin"))

# Function to plot the best fit variogram for a species in a year
# We are going to use 200 km as our cutoff: That should more or less encompass 
# stations that are one transect apart from each other (transects are separated
# by a little less than one degree latitude, which is about 110 km),
# and the extra is to include stations that are on neighboring transects
# but on opposite ends of the transect
# We will bin by every 10 km; this should get neighboring stations but not 
# stations that are two stations away
fit_variogram_one_year <- function(data, taxon, select_year){
  data_subset <- subset(data, genus_species == taxon & year == select_year)
  coordinates(data_subset) = ~X+Y
  
  # Fit isotropic variogram
  iso_v <- variogram(log(sum_of_density_number_m3+1)~1, data = data_subset, cloud=F,
                     cutoff = 200, width = 10)
  model.choices <- vgm(model = c("Nug", "Exp", "Sph", "Gau", "Lin"))
  # variogram_bestfit <- fit.variogram(iso_v, model.choices)
  
  # plot(iso_v, pl=F, model=variogram_bestfit,col="black", cex=1, lwd=0.5,lty=1,pch=20,
  #    main = paste(select_year), xlab="Distance (km)",ylab="Semivariance",
  #    xaxt = "n",
  #    xlim = c(-5, 205))
    plot(iso_v, pl=F,col="black", cex=1, lwd=0.5,lty=1,pch=20,
     main = paste(select_year), xlab="Distance (km)",ylab="Semivariance",
     xaxt = "n",
     xlim = c(-5, 205))
}

# compute Moran's I in any given year
calculate_moranI_one_year <- function(data, taxon, select_year){
  # function based on code from here: https://stats.oarc.ucla.edu/r/faq/how-can-i-calculate-morans-i-in-r/
  
  # subset data
  data_subset <- subset(data, genus_species == taxon & year == select_year)
  
  # get distances between samples
  sample.dists <- as.matrix(dist(cbind(data_subset$X, data_subset$Y)))
  
  # get inverse of matrix values and replace diagonals with zero
  sample.dists.inv <- 1/sample.dists
  diag(sample.dists.inv) <- 0
  
  # calculate Moran's I for the log density
  moran_results <- Moran.I(log(data_subset$sum_of_density_number_m3+1), sample.dists.inv)
  
  return(moran_results$p.value)
}

```


We will first examine spatial autocorrelation in *Calanus marshallae*:

```{r spatial_autocorrelation_cama, fig.show = "hold", out.width = "25%"}
### Calanus marshallae
# Figure out what years have any catch of this taxon
bongo_catch_LH_collapse %>% 
  filter(genus_species == "Calanus marshallae") %>% 
  group_by(year) %>% 
  summarise(year_total = sum(sum_of_density_number_m3)) %>% 
  filter(year_total > 0) %>% 
  arrange(year) -> cama_year_counts
```

```{r spatial_autocorrelation_cama_moran_I}
# Calculate Moran's I for each
moran_I_p_values_cama <- data.frame(year = cama_year_counts$year,
  p_value = rep(NA, length(cama_year_counts$year)))

for (i in 1:length(cama_year_counts$year)){
  moran_I_p_values_cama$p_value[i] <- calculate_moranI_one_year(data = bongo_catch_LH_collapse, 
                         taxon = "Calanus marshallae",
                         select_year = cama_year_counts$year[i])
    
}
plot(x = moran_I_p_values_cama$year, y = moran_I_p_values_cama$p_value,
     xlab = "Year", ylab = "Moran's I p-value", main = "Calanus marshallae Moran's I")
abline(a = 0.05, b = 0, lty = 2, col = "blue")
```

<br>




<br>

We will next examine spatial autocorrelation in *Calanus pacificus*:

```{r spatial_autocorrelation_capa, fig.show = "hold", out.width = "25%"}
### Calanus pacificus
# Figure out what years have any catch of this taxon
bongo_catch_LH_collapse %>% 
  filter(genus_species == "Calanus pacificus") %>% 
  group_by(year) %>% 
  summarise(year_total = sum(sum_of_density_number_m3)) %>% 
  filter(year_total > 0) %>% 
  arrange(year) -> capa_year_counts
```

```{r spatial_autocorrelation_capa_moran_I}
# Calculate Moran's I for each
moran_I_p_values_capa <- data.frame(year = capa_year_counts$year,
  p_value = rep(NA, length(capa_year_counts$year)))

for (i in 1:length(capa_year_counts$year)){
  moran_I_p_values_capa$p_value[i] <- calculate_moranI_one_year(data = bongo_catch_LH_collapse, 
                         taxon = "Calanus pacificus",
                         select_year = capa_year_counts$year[i])
    
}

plot(x = moran_I_p_values_capa$year, y = moran_I_p_values_capa$p_value,
     xlab = "Year", ylab = "Moran's I p-value", main = "Calanus pacificus Moran's I")
abline(a = 0.05, b = 0, lty = 2, col = "blue")
```

Based on the Moran's I results, we see that there is evidence for spatial clustering in most years, especially for *Calanus pacificus*.
