---
title: "JSOES Trawl Data Exploratory Analysis"
author: "Markus Min"
date: "`r Sys.Date()`"
output:
  html_document
---
This page contains exploratory analysis for the trawl data from [JSOES](https://markusmin.github.io/JSOES-SDM/jsoes_survey_description.html).

## Data Overview

We will first explore the most common and most abundant taxa in this dataset.


```{r chunk-options, message = FALSE, echo = FALSE, warning = FALSE}
# set chunk options
knitr::opts_chunk$set(message = FALSE, echo = FALSE, warning = FALSE)
```


```{r load-libraries-reformat-data}
library(readxl)
library(here)
library(janitor)
library(tidyverse)
library(kableExtra)
library(sf)
library(gstat)
library(sp)
library(ggthemes)
library(ape)

#### Species data ####
jsoes <- clean_names(read_excel(here::here("Data", "Markus_Min_Trawl_CTD_Chl_Nuts_Thorson_Scheuerell_5.2.24_FINAL.xlsx")))

species_column_names <- colnames(jsoes)[32:ncol(jsoes)]

# change nmi to km to keep units consistent
jsoes %>% 
  mutate(km_from_shore = nmi_from_shore * 1.852) -> jsoes

jsoes %>% 
  pivot_longer(., cols = all_of(species_column_names), values_to = "n_per_km", names_to = "species") %>% 
  mutate(n = n_per_km*trawl_dist_km) -> jsoes_long

# Drop any samples that don't have spatial data
filter(jsoes_long, !(is.na(mid_long))) -> jsoes_long
# this drops one two samples: 061999CR35 and 062416QR14

# turn this into an sf object
st_as_sf(jsoes_long, coords = c("mid_long", "mid_lat"), crs = 4326) -> jsoes_long_sf

# change CRS to UTM zone 10 (to work in meters)
UTM_zone_10_crs <- 32610
jsoes_long_sf_proj <- sf::st_transform(jsoes_long_sf, crs = UTM_zone_10_crs)

# make this projection into kilometers to help with interpretability
jsoes_long_sf_proj_km <- st_as_sf(jsoes_long_sf_proj$geometry/1000, crs = UTM_zone_10_crs)

# extract geometry
as.data.frame(st_coordinates(jsoes_long_sf_proj_km)) -> jsoes_long_km
# add this back to jsoes_long (X and Y now represent eastings and northings in km)
bind_cols(jsoes_long, jsoes_long_km) -> jsoes_long

# Handle repeat tows: If a tow was repeated, keep only the repeat
jsoes_long %>% 
  mutate(tow_year = paste0(station, "_", year)) -> jsoes_long
repeat_tows <- unique(subset(jsoes_long, `repeat` == TRUE)$tow_year)
jsoes_long %>% 
  filter(!(tow_year %in% repeat_tows & `repeat` == FALSE)) -> jsoes_long


# get all unique tows
jsoes_long %>% 
  dplyr::select(-c(species, n_per_km, n)) %>% 
  filter(!(duplicated(station_code))) -> jsoes_tows
```

```{r load_spatial_data, cache = TRUE}
#### Spatial data ####
usa_spdf <- st_read("/Users/markusmin/Documents/ESA_RF_2021/map_files/USA_adm0.shp", quiet = TRUE)
# load BC
CAN_spdf <- st_read("/Users/markusmin/Documents/ESA_RF_2021/map_files/canada/lpr_000b16a_e.shp", quiet = TRUE)
BC_spdf <- filter(CAN_spdf, PRENAME == "British Columbia")
BC_proj <- st_transform(BC_spdf, crs = 4326)


# crop them to our desired area
US_west_coast <- sf::st_crop(usa_spdf,
    c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))

BC_coast <- sf::st_crop(BC_proj,
    c(xmin = -126, ymin = 44, xmax = -123, ymax = 48.5))



# convert both shapefiles to a different projection (UTM zone 10) so that they can be plotted with the sdmTMB output
UTM_zone_10_crs <- 32610

US_west_coast_proj <- sf::st_transform(US_west_coast, crs = UTM_zone_10_crs)
BC_coast_proj <- sf::st_transform(BC_coast, crs = UTM_zone_10_crs)

# make this projection into kilometers
US_west_coast_proj_km <- st_as_sf(US_west_coast_proj$geometry/1000, crs = UTM_zone_10_crs)
BC_coast_proj_km <- st_as_sf(BC_coast_proj$geometry/1000, crs = UTM_zone_10_crs)
```

```{r create_basemap, cache = TRUE}
# create base map
survey_area_basemap_km <- ggplot(US_west_coast_proj_km) +
  geom_sf() +
  geom_sf(data = BC_coast_proj_km) +
  # coord_sf(ylim = c(4922.052, 5342.052), xlim = c(334.8638, 404.8638)) +
  # coord_sf(ylim = c(44,48.5),  xlim = c(-126, -123)) +
  ylab("Latitude")+
  xlab("Longitude")+
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white", color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.14, 0.2),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12),
        axis.ticks = element_blank(),
        axis.text = element_blank())
  # temporary fix for lat/long on axes to just get rid of them - start here to actually fix: https://forum.posit.co/t/converting-axes-to-lat-lon/27181

survey_area_basemap <- ggplot(US_west_coast) +
  geom_sf() +
  geom_sf(data = BC_coast) +
  coord_sf(ylim = c(44,48.5),  xlim = c(-126, -123)) +
  scale_x_continuous(breaks = c(124,125,126)) +
  ylab("Latitude")+
  xlab("Longitude")+
  theme(plot.background = element_rect(fill = "white"),
        panel.background = element_rect(fill="white", color = "black"),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.14, 0.2),
        legend.title = element_text(size = 14),
        legend.text = element_text(size = 12))
```

```{r common_taxa}
## total density
jsoes_long %>% 
  group_by(species) %>% 
  summarise(mean_n_per_km = mean(n_per_km),
            sd_n_per_km = sd(n_per_km)) %>% 
  arrange(desc(mean_n_per_km)) -> jsoes_density_sums

jsoes_density_sums[1:10,] %>% 
  mutate(mean_n_per_km = round(mean_n_per_km, 1),
         sd_n_per_km = round(sd_n_per_km, 1)) %>% 
  kbl(caption = "Top ten taxa by mean density.") %>% 
  kable_styling(latex_options = "HOLD_position")

## frequency of occurrence

# get count of each genus_species
jsoes_long %>% 
  filter(n > 0) %>% 
  count(species) %>% 
  arrange(desc(n)) %>% 
  mutate(prop_samples = n/nrow(jsoes)) -> jsoes_freq_occ

jsoes_freq_occ[1:10, ] %>% 
  mutate(prop_samples = round(prop_samples, 2)) %>% 
  kbl(caption = "Top ten taxa by frequency of occurrence.") %>% 
  kable_styling(latex_options = "HOLD_position")

```

## Annual time series

This Shiny app can be used to explore the abundances of different taxa across the full length of the time series. In this Shiny app, I take the mean log density across the survey region to create a simple index of abundance.

```{r}
knitr::include_app(url = "https://markusmin.shinyapps.io/jsoes_trawl_annual_time_series/", height = "700px")
# <iframe src="https://markusmin.shinyapps.io/jsoes_trawl_annual_time_series/" style="border:none;width:1000px;height:500px;"></iframe>
```


## What salmon are we getting?

By frequency of occurrence:
- Coho salmon yearling (coho_salmon_yearling) - 55%
  - My understanding is that if these are Columbia River Coho, they're most likely from downstream of BON
  and therefore aren't good candidates for SAR estimates
- Chinook yearling Interior Fall (chinook_salmon_yearling_interior_fa) - 33%
  - Should double check where these are coming from
- Chum salmon juvenile (chum_salmon_juvenile) - 30%
  - I don't think people really care about Chum. Sorry Chum.
- Chinook yearling Interior Spring (chinook_salmon_yearling_interior_sp) - 25%
  - definitely include
- Chinook salmon mixed age juvenile (chinook_salmon_mixed_age_juvenile) - 25%
  - unknown origin?
- Chinook subyearling Interior Fall (chinook_salmon_subyearling_interior_fa) - 24%
  - definitely include
- Chinook salmon yearling West Cascade spring (lower Columbia) (chinook_salmon_yearling_wc_sp) - 13%
  - Lower Columbia - likely downstream of BON
- Chinook salmon yearling Willamette River Spring (chinook_salmon_yearling_wr_sp) - 13%
  - Willamette River ESU, downstream of BON
- Sockeye salmon juvenile (sockeye_salmon_juvenile) - 12%
  - unknown origin
- Chinook Spring Creek Group fall (chinook_salmon_subyearling_scg_f) - 8%
  - Lower Columbia - likely downstream of BON
- Chinook salmon Willamette River Subyearling Spring (chinook_salmon_subyearling_wr_sp) - 8%
  - Willamette River ESU, downstream of BON

We need to understand what GSI groups are even worth modeling SAR data for
```{r}
# 5% cutoff?
subset(jsoes_freq_occ, grepl("salmon", species))
subset(jsoes_freq_occ, grepl("salmon", species) & prop_samples > 0.05)
```





## Plotting distributions of some common taxa

I am currently developing a Shiny app that will allow you to visualize the distributions of different taxa. For now, I provide static maps of four focal taxa for this survey:

- Yearling Coho Salmon
- Yearling Interior Columbia Fall Chinook Salmon
- Yearling Interior Columbia Spring Chinook Salmon
- Water Jellies

```{r prep_salmon_data}
csyis <- subset(jsoes_long, species == "chinook_salmon_yearling_interior_sp")
coho <- subset(jsoes_long, species == "coho_salmon_yearling")
csyif <- subset(jsoes_long, species == "chinook_salmon_yearling_interior_fa")
cssif <- subset(jsoes_long, species == "chinook_salmon_subyearling_interior_fa")
water_jellies <- subset(jsoes_long, species == "water_jelly")
```



```{r map_salmon, cache = TRUE}
plot_distribution <- function(data, taxon_name){
  # create facet_wrap plot for distribution across all years
  data %>% 
    mutate(encounter = ifelse(n_per_km == 0, "zero", "non-zero")) -> data
  
  # Drop 1998 (no water jellies recorded in that year)
  data %>% 
    filter(year != 1998) -> data
  
survey_area_basemap +
      geom_point(data = data, aes(x = mid_long, y = mid_lat, size = n_per_km, color = encounter),
                 alpha = 0.5) +
    scale_color_manual(values = c("zero" = "#fc9272", "non-zero" = "#2ca25f")) +
    facet_wrap(~year, nrow = 3) +
    ggtitle(taxon_name) +
    theme(legend.position = c(0.9, 0.15),
        legend.key.height = unit(0.35, "cm"),
        legend.key.width = unit(0.25, "cm"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        legend.spacing.y = unit(0.01, 'cm')) -> species_distribution_plot
  
  return(species_distribution_plot)
}

coho_dist_plot <- plot_distribution(data = coho, taxon_name = "Yearling Coho Salmon")
csyis_dist_plot <- plot_distribution(data = csyis, taxon_name = "Yearling Interior Columbia Spring Chinook")
csyif_dist_plot <- plot_distribution(data = csyif, taxon_name = "Yearling Interior Columbia Fall Chinook")
cssif_dist_plot <- plot_distribution(data = cssif, taxon_name = "Subyearling Interior Columbia Fall Chinook")
water_jellies_dist_plot <- plot_distribution(data = water_jellies, taxon_name = "Water Jellies")

coho_dist_plot
csyis_dist_plot
csyif_dist_plot
cssif_dist_plot
water_jellies_dist_plot

ggsave(here::here("figures", "csyis_distribution_plot.png"), plot = csyis_dist_plot, height = 12, width = 17)
ggsave(here::here("figures", "csyif_distribution_plot.png"), plot = csyif_dist_plot, height = 12, width = 17)
ggsave(here::here("figures", "cssif_distribution_plot.png"), plot = cssif_dist_plot, height = 12, width = 17)
ggsave(here::here("figures", "coho_distribution_plot.png"), plot = coho_dist_plot, height = 12, width = 17)
ggsave(here::here("figures", "water_jellies_distribution_plot.png"), plot = water_jellies_dist_plot, height = 12, width = 17)

```


## Temporal and Spatial Autocorrelation

Before fitting any spatiotemporal models, we must explore the spatial and temporal autocorrelation in the data.

<br>

### Temporal structure

We can first inspect the autocorrelation in our four focal taxa, at the scale of the full survey region:

```{r coastwide_temporal_autocorrelation}
create_time_series <- function(data, taxon){
  data %>% 
    filter(year != 1998)  %>% 
    mutate(log_density = log(n_per_km + 1)) %>% 
    group_by(year) %>% 
    summarise(mean_density = mean(log_density),
              sd_density = sd(log_density)) %>% 
    # mutate(CI_95_upper = mean_density + 1.96*sd_density) %>% 
    # mutate(CI_95_lower = mean_density - 1.96*sd_density) %>% 
    mutate(taxon = taxon) -> ts
  
  return(ts)
}

create_time_series(data = csyis,
                   taxon = "Interior Columbia Yearling Spring Chinook") -> csyis_ts
create_time_series(data = csyif,
                   taxon = "Interior Columbia Yearling Fall Chinook") -> csyif_ts
create_time_series(data = coho,
                   taxon = "Yearling Coho Salmon") -> coho_ts
create_time_series(data = water_jellies,
                   taxon = "Water Jellies") -> water_jellies_ts
# It would make sense to try different spatial structures and then test autocorrelation.
# 1) Coastwide; 2) By transect; 3) By station

acf(csyis_ts$mean_density, main = "Interior Columbia Yearling Spring Chinook")
acf(csyif_ts$mean_density, main = "Interior Columbia Yearling Fall Chinook")
acf(coho_ts$mean_density, main = "Yearling Coho Salmon")
acf(water_jellies_ts$mean_density, main = "Water Jellies")
    
```

None of the four taxa investigated here show temporal autocorrelation.

I also investigated temporal autocorrelation at the scale of individual stations, and found little evidence for temporal autocorrelation at this scale for these four taxa.


<br>

### Spatial structure

To investigate spatial autocorrelation, I calculated a metric known as Moran's I. Moran's I is a measure of the overall clustering of the spatial data and tests if there is support to reject the null hypothesis of no spatial structure. Given that we did not see much evidence for temporal autocorrelation, I calculated Moran's I separately for each year. To summarize the Moran's I results, I show the p-value for Moran's I for each year, with the blue dashed line showing a p-value of 0.05.

```{r spatial_autocorrelation_functions}
# If you wanted to add a best fit line:
# A best fit line is then plotted through the points.
# For each year, we will plot the best fit semivariogram model, which we selected from a set of five candidate functional response shapes: 1) Exponential, 2) Spherical, 3) Gaussian, 4) Linear, and 5) Nugget only (y-intercept only).



# Set the models to test
# To see the shape of these models, run `show.vgms()`
model.choices <- vgm(model = c("Nug", "Exp", "Sph", "Gau", "Lin"))

# Function to plot the best fit variogram for a species in a year
# We are going to use 200 km as our cutoff: That should more or less encompass 
# stations that are one transect apart from each other (transects are separated
# by a little less than one degree latitude, which is about 110 km),
# and the extra is to include stations that are on neighboring transects
# but on opposite ends of the transect
# We will bin by every 10 km; this should get neighboring stations but not 
# stations that are two stations away
fit_variogram_one_year <- function(data,  select_year){
  data_subset <- subset(data, year == select_year)
  coordinates(data_subset) = ~X+Y
  
  # Fit isotropic variogram
  iso_v <- variogram(log(n_per_km+1)~1, data = data_subset, cloud=F,
                     cutoff = 200, width = 10)
  model.choices <- vgm(model = c("Nug", "Exp", "Sph", "Gau", "Lin"))
  variogram_bestfit <- fit.variogram(iso_v, model.choices)
  
  # plot(iso_v, pl=F, model=variogram_bestfit,col="black", cex=1, lwd=0.5,lty=1,pch=20,
  #    main = paste(select_year), xlab="Distance (km)",ylab="Semivariance",
  #    xaxt = "n",
  #    xlim = c(-5, 205))
    plot(iso_v, pl=F,col="black", cex=1, lwd=0.5,lty=1,pch=20,
     main = paste(select_year), xlab="Distance (km)",ylab="Semivariance",
     xaxt = "n",
     xlim = c(-5, 205))
}

# compute Moran's I in any given year
calculate_moranI_one_year <- function(data, select_year){
  # function based on code from here: https://stats.oarc.ucla.edu/r/faq/how-can-i-calculate-morans-i-in-r/
  
  # subset data
  data_subset <- subset(data, year == select_year)
  
  # get distances between samples
  sample.dists <- as.matrix(dist(cbind(data_subset$X, data_subset$Y)))
  
  # get inverse of matrix values and replace diagonals with zero
  sample.dists.inv <- 1/sample.dists
  diag(sample.dists.inv) <- 0
  
  # calculate Moran's I for the log density
  moran_results <- Moran.I(log(data_subset$n_per_km+1), sample.dists.inv)
  
  return(moran_results$p.value)
}

```

<br>


<br>

```{r spatial_autocorrelation_coho, fig.show = "hold", out.width = "25%"}
### Yearling Coho Salmon
# Figure out what years have any catch of this taxon
coho %>% 
  group_by(year) %>% 
  summarise(year_total = sum(n_per_km)) %>% 
  filter(year_total > 0) %>% 
  arrange(year) -> coho_year_counts
```

```{r spatial_autocorrelation_coho_moran_I}
# Calculate Moran's I for each
moran_I_p_values_coho <- data.frame(year = coho_year_counts$year,
  p_value = rep(NA, length(coho_year_counts$year)))

for (i in 1:length(coho_year_counts$year)){
  moran_I_p_values_coho$p_value[i] <- calculate_moranI_one_year(data = coho,
                         select_year = coho_year_counts$year[i])
    
}
plot(x = moran_I_p_values_coho$year, y = moran_I_p_values_coho$p_value,
     xlab = "Year", ylab = "Moran's I p-value", main = "Yearling Coho Salmon Moran's I")
abline(a = 0.05, b = 0, lty = 2, col = "blue")
```

<br>


<br>

```{r spatial_autocorrelation_csyis, fig.show = "hold", out.width = "25%"}
### Yearling Interior Spring Chinook Salmon
# Figure out what years have any catch of this taxon
csyis %>% 
  group_by(year) %>% 
  summarise(year_total = sum(n_per_km)) %>% 
  filter(year_total > 0) %>% 
  arrange(year) -> csyis_year_counts
```

```{r spatial_autocorrelation_csyis_moran_I}
# Calculate Moran's I for each
moran_I_p_values_csyis <- data.frame(year = csyis_year_counts$year,
                                    p_value = rep(NA, length(csyis_year_counts$year)))

for (i in 1:length(csyis_year_counts$year)){
  moran_I_p_values_csyis$p_value[i] <- calculate_moranI_one_year(data = csyis,
                                                                select_year = csyis_year_counts$year[i])
  
}
plot(x = moran_I_p_values_csyis$year, y = moran_I_p_values_csyis$p_value,
     xlab = "Year", ylab = "Moran's I p-value", main = "Yearling Interior Spring Chinook Salmon Moran's I")
abline(a = 0.05, b = 0, lty = 2, col = "blue")
```

<br>


<br>


```{r spatial_autocorrelation_csyif, fig.show = "hold", out.width = "25%"}
### Yearling Interior Fall Chinook Salmon
# Figure out what years have any catch of this taxon
csyif %>% 
  group_by(year) %>% 
  summarise(year_total = sum(n_per_km)) %>% 
  filter(year_total > 0) %>% 
  arrange(year) -> csyif_year_counts

```

```{r spatial_autocorrelation_csyif_moran_I}
# Calculate Moran's I for each
moran_I_p_values_csyif <- data.frame(year = csyif_year_counts$year,
                                    p_value = rep(NA, length(csyif_year_counts$year)))

for (i in 1:length(csyif_year_counts$year)){
  moran_I_p_values_csyif$p_value[i] <- calculate_moranI_one_year(data = csyif,
                                                                select_year = csyif_year_counts$year[i])
  
}
plot(x = moran_I_p_values_csyif$year, y = moran_I_p_values_csyif$p_value,
     xlab = "Year", ylab = "Moran's I p-value", main = "Yearling Interior Fall Chinook Salmon Moran's I")
abline(a = 0.05, b = 0, lty = 2, col = "blue")
```

<br>


<br>


```{r spatial_autocorrelation_water_jellies, fig.show = "hold", out.width = "25%"}
### Water Jellies
# Figure out what years have any catch of this taxon
water_jellies %>% 
  group_by(year) %>% 
  summarise(year_total = sum(n_per_km)) %>% 
  filter(year_total > 0) %>% 
  arrange(year) -> water_jellies_year_counts

```

```{r spatial_autocorrelation_water_jellies_moran_I}
# Calculate Moran's I for each
moran_I_p_values_water_jellies <- data.frame(year = water_jellies_year_counts$year,
                                    p_value = rep(NA, length(water_jellies_year_counts$year)))

for (i in 1:length(water_jellies_year_counts$year)){
  moran_I_p_values_water_jellies$p_value[i] <- calculate_moranI_one_year(data = water_jellies,
                                                                select_year = water_jellies_year_counts$year[i])
  
}
plot(x = moran_I_p_values_water_jellies$year, y = moran_I_p_values_water_jellies$p_value,
     xlab = "Year", ylab = "Moran's I p-value", main = "Water Jellies Moran's I")
abline(a = 0.05, b = 0, lty = 2, col = "blue")
```

Based on the Moran's I results, we see that there is evidence for spatial clustering in some years but not others. There are also notably more years where the Moran's I value is significant for these taxa than there are for the two copepod species examined in the JSOES Bongo exploratory analysis.

